\thispagestyle{empty}
\vspace*{3cm}

\begin{center}
\Large
\textbf{Dedicated to}

\vspace{2em}

\textit{Those in my life who made me feel special—not because they thought I was special, but because they introduced me to a world of what seemed like magic at the time.}

\vspace{2em}

\textbf{Norman Zocher}\\
\normalsize
Who introduced me one day to his mother, before a performance, as a Blues guitarist prodigy. I don't know if he thought I was for real. But God, I never believed it—and it did so much for me to this day. I'm glad he's still alive, and I hope he reads this one day. Thank you, Norman.

\vspace{1.5em}

\Large
\textbf{Henry Tate}\\
\normalsize
The most wonderful person, teacher, and friend. Henry didn't teach art. He taught life. I am amongst many—some very famous—who are in complete agreement about this remarkable man.

\vspace{1.5em}

\Large
\textbf{Tibor Pusztai}\\
\normalsize
He taught me how to count polyrhythms by explaining it through the analogy of how artillery fire uses tracers every seven rounds. He was a wonderful man, wonderful composer. The world of polyrhythms, the whimsy of Bartók—not just musically, but thematically. To this day I am embarrassed to say I don't particularly enjoy opera, but he made me appreciate it in a way that nobody else could.

\vspace{1.5em}

\Large
\textbf{Juan José Calatayud}\\
\normalsize
My piano teacher for probably three lessons in Mexico, because of a favor from my grandmother. Probably three hours of my life. They mattered.

\vspace{1.5em}

\Large
\textbf{Marcos Kurtycz}\\
\normalsize
Growing up as a child around the greatest of artists, Marcos inspired not only my work scientifically, but as a painter and musician. His spirit continues in the same legacy as Tibor's. They even kind of look the same, with a pronounced height distinction.

\vspace{2em}

\normalsize
\textit{If you ever wondered why this book exists, it was because of superhuman teachers like these men. Who not only made me feel normal—they kept me engaged. They taught me things so valuable that almost 30 years after the fact, I am crying as I type these words. Out of happiness that they were in my life.}
\end{center}

\clearpage
